# Graphics with ggplot2

## Basic knowledge of ggplot2 
There are two ways to produce plots in ggplot2, one is `qplot()` and the other is `ggplot()`. In this seminar we focus on `ggplot`.
  
### Basic Components of a ggplot2 Plot





[Data Visualization with ggplot2 Cheat Sheet - RStudio](https://rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf)


### Building a simple ggplot plot
Now let us start with creating a scatter plot using ggplot().The scatter plot is created to exploring the relationship between transaction price and total floor area in two different local authorities.  
#### step 1 plot background  
```{r}
ggplot(housedata1,aes(x=tfarea,y=price))
```
#### step 2 add a layer

```{r}
ggplot(housedata1,aes(x=tfarea,y=price))+
  geom_point()
```

#### step 3  add the thid layer(linear regression line)

```{r}
ggplot(housedata1,aes(x=tfarea,y=price))+
  geom_point()+
  geom_smooth(method="lm")
```



#### step 4 add anoher layer
```{r}
ggplot(housedata1,aes(x=tfarea,y=price))+
  geom_point()+
  geom_smooth(method="lm")+
  facet_wrap(~ldnm)
```
#### step 4  add a linear regression line


####  

###
####


####

####



### Extensions to ggplot2

ggthemes – more diagram themes, allows you to make graphics appear as if produced by Stata,
Excel, The Economist, etc.
ggformula – an interface similar to lattice
ggpmisc – support for time series and polynomial ts
ggtern – ternary diagrams for compositional data
ggforce – various “geom” layers, including pie charts, and other extensions
ggExtra – add marginal plots etc
 GGally – special purpose plots, coecient plots, model diagnostics, networks, correlation matrices
 plotly has the function ggplotly to translate “ggplot” results into JavaScript diagrams suitable for
websites.



Since ggplot only works with **data** frames, 


```{r}
#test the data set is a data frame
is.data.frame(housedata1)
is.data.frame(housedata2)
```





### 


####
Here is a review of existing methods.





## Sub and superscript



## Saving graphs
TIFF files exported directly from RStudio will only achieve a resolution of 72dpi.


In RStudio, you can click on the “Export” button and export your plots


Using ggsave is used to save a ggplot2 graph from the screen to a file

```{r}
ggsave("Figure_B.tiff",units="in", width=14, height=6, dpi=500, compression = 'lzw')  
```

## Sum up



